---
title: "DATA 607 Project 2"
author: "Curwen Arthurs"
date: "March 12, 2017"
output: html_document
---

## Working with wide and untidy data sets

# What is a wide dataset?

Wide, or unstacked data is presented with each different data variable in a separate column. "https://en.wikipedia.org/wiki/Wide_and_narrow_data"

 

# What is untidy data?

Untidy data is messy or poorly formatted which makes is harder to do analysis.


# Tidy Rules

1. Each variable in the dataset is placed in its own column

2. Each observation is placed in its own row

3. Each value is placed in its own cell

source: http://garrettgman.github.io/tidying/


## Load libraries
```{r}
library(tidyr)
library(dplyr)
library(zoo)
library(stringr)
```


## Dataset 1 

# Read in the csv file into R

The file can be downloaded via this link:

https://bbhosted.cuny.edu/courses/1/SPS01_DATA_607_01_1172_1/db/_22605542_1/Discussion5.csv



## Loading Dataset 1   

```{r}
#This code reads the Discussion5.csv file into data1 object.


data1<- read.csv("C:/CUNYSPS/Spring 2017/Data Acquisition Management/Week6/Discussion5.csv", na.strings=c("","NA"), stringsAsFactors = FALSE) 
# An additional argument is added to label all blank rows as "NA" in order to use the na.locf function
```

# Start Tidying 

```{r}
##1
# First we need to repeat the date for each Carrier item. Thats is for each Carrier repeating the date for the shipping fee.

# By using the mutate function we can update all rows missing a date value that we earlier labeled "NA" with the previous non missing value. We do this by using the na.locf function.

data1.1<- mutate(data1, Date = na.locf(Date))


##2
#Second need to make the dataset into a long instead of wide format. This is done using the gather function to convert columns into rows.

##3
#Third we need to ensure that each column only contains 1 variable. Notice that the values column has both price and shipping fees so we need to change that to make the dataset tidy. We can split these into 2 different columns by using the spread function.

# Lastly we can order the dataset by country in alphabetical order using the arrange function.


data1.2<-
  data1.1 %>%
  gather("Country", "Amount", 3:10) %>% 
  spread(Values, Amount) %>%
  arrange(Country)
```

## Dataset 1 - Analysis

```{r}
#Analysis 1 - Percentage of tax charged 
#Question: are we collecting enough fees vs what the company is charging

analysis1<- data1.2 %>%
            mutate(Percentage = Price of Carrier)

#Analysis 2 - Average shipping cost per month

analysis<- data1.2 %>%

#Analysis 3 - Average shipping cost per month by country

```




## Dataset 2 

Table was copied from https://www.bls.gov/news.release/empsit.t19.htm and pasted into an excel sheet. It was then saved as a comma separated file "data2"


# Reading the the dataset into R and extract just the table results

```{r}
# Create an object data2 and store the information from the csv file in it
# Remove table title information by starting reading the from line 3 of the csv file. Set parameter skip = 2.
# Set all blank values to NA using na.strings argument

data2<- read.csv("C:/CUNYSPS/Spring 2017/Data Acquisition Management/Week6/data2.csv", skip = 2, stringsAsFactors = FALSE)

# Remove footnotes and note information which is not relevant to the table for analysis.
#Can do this by using the slice function to subset only those rows we want.

data2<- slice(data2, 1:21)
```


## Start Tidying data
# Split table into 2

```{r}
#First create a new table for Average hourly earnings by selecting the columns for that using the select function then save it as its own table 
data2.1<- select(data2, Industry, Average.hourly.earnings, X, X.1, X.2)


# Do the same process to create a new table for Average weekly earnings
data2.2<- select(data2, -Average.hourly.earnings, -X, -X.1, -X.2)
```




## Tidy 2 tables separately
# Tidy hourly table

```{r}
#rename the column headings combining the 2 separate values of month and year
colnames(data2.1) <- c("Industry", "Feb 2016", "Dec 2016", "Jan 2017(p)", "Feb 2017(p)")

data2hrly<-
           data2.1 %>%
           slice(3:21) %>% #removes the 2 rows contained the column heading info
           gather("Date", "Average hourly earnings in dollars", 2:5) #convert columns into rows for tidyness

```

# Tidy weekly table

```{r}
#rename the column headings combining the 2 separate values of month and year
colnames(data2.2) <- c("Industry", "Feb 2016", "Dec 2016", "Jan 2017(p)", "Feb 2017(p)")

data2wkly<-
           data2.2 %>%
           slice(3:21) %>% #removes the 2 rows contained the column heading info
           gather("Date", "Average weekly earnings in dollars", 2:5) #convert columns into rows for tidyness

```


## Merge 2 tidy datasets into 1 final tidy dataset
```{r}
#Merge data2hrly with data2wkly using inner join and with a composite key where the industry type and date make up the composite key
data2final<- inner_join(data2hrly,data2wkly, by = c("Industry" = "Industry", "Date" ="Date"))
              
data2final<- mutate_each(data2final,funs(replace(., .="$", "")), -Industry, -Date)

data2final1<- mutate(data2final, ))
data2final1<- mutate(data2final, str_sub(data2final[,3:4], 1, str_length(data2final[,3:4])==1))

# Remove "$" from total values in Average hourly and weekly earnings columns
data.frame(lapply(data2final, gsub, pattern = "$", replacement = " "))
```


## Analysis
```{r}


```








