---
title: "Assignment - Web APIs"
author: "Curwen Arthurs"
date: "April 2, 2017"
output: html_document
---

## Loading Libraries to use
```{r}
library(httr)
library(jsonlite)
```


## Exploring the New York Times API

Anyone wishing to call a NYTimes API must register and requset an API key.

http://developer.nytimes.com/docs

The Times offers APIs in the following categories:

- Archive

- Article Search

- Books

- Community

- Geographic

- Most Popular

- Movie Reviews

- Semantic

- Times Newswire

- TimesTags

- Top Stories


After browsing through each of the categories I decided to use the Article Search API category.

## ABOUT THE API

The NYT Article Search API allows you to search more than 2.8 million New York Times articles from 1981 to today, retrieving headlines, abstracts, lead paragraphs and links to associated multimedia.

The API supports the following type of searching: 
- Standard keyword searching
- Date range: all articles from X date to Y date
- Field search: search within any number of given fields, e.g., title:obama byline:dowd
- Conjunction and disjunction (AND and NOT) operations, e.g., baseball yankees -"red sox"
- Ordering by closest (variable ranking algorithms), newest and oldest



## Load the API

#Parameters

```{r, include = false}
apikey<- c("63c7bd12f65341b38d7d810a87973ae1")
baseurl<- "https://api.nytimes.com/svc/search/v2/articlesearch.json?"
terms<- c("Terrorist+Attacks")
bdate<- c("19980101")
edate<- c("20010910")
```


## Create url

```{r}
nytsearch.url<- paste0(baseurl, "q=", terms, "&", bdate, "&", edate, "&", "api-key=", apikey)

nytsearch.url
```



## Parse the Articles

```{r}

#store all pages in a list first
pages <- list()

#create a basic loop to store the results of each page call, for a total of 94 
for(i in 0:93){
  mydata <- fromJSON(paste0(nytsearch.url, "&page=", i))
  message("Retrieving page ", i)
  pages[[i+1]] <- mydata$response
  Sys.sleep(1) # sleep for 1 second in between calls to avoid HTTP 429 error
}

#combine all into one
#response<- rbind.pages(pages[sapply(pages, length)>0])
```


## Convert to data frame

```{r}
nytsearchdf<- as.data.frame(pages)
```

