---
title: "R Character Manipulation and Date Processing"
author: "Curwen Arthurs"
date: "February 19, 2017"
output: html_document
---

## Load libraries to use.

```{r}
library(stringr)
library(RCurl)
library(XML)
```

## Create string vector containing phone numbers and names.

```{r}
raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"

name<- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
```

## Create subset of this vector that has only names and call it name.

3. (a) Use the tools of this chapter to rearrange the vector so that all elements conform to
the standard first_name last_name.

```{r}
char2<- str_replace(name[[2]], "([[:alpha:]]{5}.) ([[:alpha:]]{1}.) ([[:alpha:]]{10})" , "\\3 \\2 \\1")
char5<- str_replace(name[[5]], "([[:alpha:]]{7}.) ([[:alpha:]]{5})" , "\\2 \\1")
name2<- str_replace(c(name[[1]], char2, name[[3]], name[[4]], char5, name[[6]]), ",", "")
name2
```


3. (b) Construct a logical vector indicating whether a character has a title (i.e., Rev. and
Dr.).

```{r}
str_detect(name, "[[:alpha:]]{2,}\\.")
```

3. (c) Construct a logical vector indicating whether a character has a second name.

```{r}
str_detect(name, "[[:alpha:]]{4}.+[[:alpha:]].+[[:alpha:]]{10}")
```




4. (a) [0-9]+\\$ - this pattern is for any amount of digits between 0-9 followed by the $ sign. Can be used to subset currency information for dollar currencies.  

```{r}
str_view_all(c("10£","100$BZ", "1000\u20AC","1000$US"), "[0-9]+\\$")
```

4. (b) \\b[a-z]{1,4}\\b - this pattern matches lower case words with of length 1 to 4 characters.

```{r}
str_view_all(c("Copy the introductory example. The vector name stores the extracted names"), "\\b[a-z]{1,4}\\b", match = TRUE)
```

4. (c) .*?\\.txt$ - this patern looks at strings that end with ".txt". Can be used to subset text files.  

```{r}
str_view_all(c("name.rmd","name.doc", "name.xlsx","name.txt",".txt"), ".*?\\.txt$")
```

4. (d) \\d{2}/\\d{2}/\\d{4} - this pattern matches dates in the formatt mm/dd/yyyy or dd/mm/yyyy

```{r}
str_view_all(c("19/02/2017","19th February, 2017","19-02-2017", "Feb.19.2017","02/19/2017 10:30:15"), "\\d{2}/\\d{2}/\\d{4}")
```

4. (e) <(.+?)>.+?</\\1> - this pattern matches html tags and their contents.

```{r}
webpage <- getURL("http://global.nba.com/wp-content/include/global/splash_page/promo.html?gr=www", followlocation=TRUE)
unlist(str_extract_all(webpage, "<(.+?)>.+?</\\1>"))
```

9. The following code hides a secret message. Crack it with R and regular expressions.

```{r}
# Source "https://www.r-bloggers.com/htmltotext-extracting-text-from-html-via-xpath/"
# download html
html <- getURL("http://www.r-datacollection.com/materials/regex/code_exercise.txt", followlocation = TRUE)
 
# parse html
doc = htmlParse(html, asText=TRUE)
code_exercise <- xpathSApply(doc, "//p", xmlValue)
cat(paste(code_exercise, collapse = "\n"))

#this code extracts all the capital letters, . and ! from the code_exercise text
#it then concatenates all the elements together into 1 string and finally replaces the "." separators with blank spaces  
str_replace_all(str_c(unlist(str_extract_all(code_exercise, "[A-Z]|\\.|!")), collapse = ""), "\\.", " ")
```


